---
title: "Exploratory Analysis"
author: "John Ho, Lucy Lin, Brianna Pak, Devina Tavathia, Kelly Wang"
date: "11/1/2021"
output: html_document
---
Libraries Used:
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(tidyverse)
library(readr)
```

## Introductory Paragraph
Our group wanted to analyze the trends and correlations, if any, from homicide cases recorded in three regional scopes: nationally 
in the U.S., in Washington state, and in the city Seattle, Washington. The data came from 
[Kaggle](https://www.kaggle.com/murderaccountability/homicide-reports?select=database.csv), which features 24 separate 
columns as an overview of details and 638454 cases.

* The Kaggle data was collected by the FBI’s Supplementary Homicide Report and Freedom of Information Act and compiled by 
the Murder Accountability Project. It is stated that the database contains homicides dating back to 1980 plus the 22,000 
not reported to the Justice Department, with classifications on race, sex, and age of both perpetrators and victims.

We wanted to explore the information through filtering data for details we would need. Things such as categorizations of 
the perpetrator and victim, weapons used, areas with higher homicide rates, span of years and number of murders happening, 
all of this information was used to answer our research queries.

### Summary Information
Here are some questions we thought of when planning our project:

* How do factors of the perpetrator and victim (ex. age, race, etc.) correlate?
* Are there trends between perpetrator age and homicide type?
* What are the comparisons of homicide rates between other states?
* Which cities of each state have the most homicide cases? Any concentrations?
* How do current homicide rates within this decade compare to rates from previous decades (ex. 2000-2018, 1990-2000)?

These questions branched into more and more questions as we sorted through our data. Some statistics we sought out for 
were:

* Cities with highest **and** lowest numbers of homicide cases in each state
* Largest number of homicides in a single state
* Average age of murder victims in Washington State
* Number of total cases (rows) altogether

#### Summary Table
The second set of questions were the questions we wanted to experiment with our data set. We decided to summarise the answers in our __Summary_Information.R__ file, which is displayed underneath, answering the questions in order (rendered in comments).

```{r echo=FALSE message=FALSE}
source("Summary_Information.R", local = knitr::knit_global())
summary_list
```

#### Homicide Counts Over the Decades


```{r echo=FALSE}
source("Scatterplot.R", local = knitr::knit_global())
year_incident_plot
```

#### Popular Weapons Used in Homicide
We wanted to investigate if there was a more commonly used weapon within all homicide cases.

```{r echo=FALSE}
source("barchart.r", local = knitr::knit_global())
weapon_barchart
```