---
title: "Exploratory Analysis"
author: "John Ho, Lucy Lin, Brianna Pak, Devina Tavathia, Kelly Wang"
date: "11/1/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(tidyverse)
library(readr)
```

## Introductory Paragraph
Our group wanted to analyze the trends and correlations, if any, from homicide cases recorded in three regional scopes: nationally 
in the U.S., in Washington state, and in the city Seattle, Washington. The data came from 
[Kaggle](https://www.kaggle.com/murderaccountability/homicide-reports?select=database.csv), which features 24 separate columns as an overview of details and 638454 cases.

* The Kaggle data was collected by the FBI’s Supplementary Homicide Report and Freedom of Information Act and compiled by the Murder 
Accountability Project. It is stated that the database contains homicides dating back to 1976 plus the 22,000 not reported to the Justice 
Department, with classifications on race, sex, and age of both perpetrators and victims.

We wanted to explore the information through filtering data for details we would need. Things such as categorizations of the perpetrator and 
victim, weapons used, areas with higher homicide rates, span of years and number of murders happening, all of this information was used to answer
our research queries.

### Summary Information
Here are some questions we thought of when planning our project:

* How do factors of the perpetrator and victim (ex. age, race, etc.) correlate?
* Are there trends between perpetrator age and homicide type?
* What are the comparisons of homicide rates between other states?
* Which cities of each state have the most homicide cases? Any concentrations?
* How do current homicide rates within this decade compare to rates from previous decades (ex. 2000-2018, 1990-2000)?

These questions branched into more and more questions as we sorted through our data. Some statistics we sought out for were:

* Cities with highest **and** lowest numbers of homicide cases in each state
* Largest number of homicides in a single state
* Average age of murder victims in Washington State
* Number of total cases (rows) altogether

#### Summary Table
```{r}
#source("Summary_Information.R", local = knitr::knit_global())
```
### Scatter Plot (xyz)


```{r echo=FALSE}
#source("Scatterplot.R", local = knitr::knit_global())
```
### Bar Graph (xyz)


```{r echo=FALSE}
#source("barchart.R", local = knitr::knit_global())
```