---
title: "Exploratory Analysis"
author: "John Ho, Lucy Lin, Brianna Pak, Devina Tavathia, Kelly Wang"
date: "11/1/2021"
output: html_document
---
Libraries Used:
```
library(dplyr)
library(ggplot2)
library(gridExtra)
library(readr)
library(tidyverse)
```
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(gridExtra)
library(readr)
library(tidyverse)
```

## Introductory Paragraph
Our group wanted to analyze the trends and correlations, if any, from homicide cases recorded in three regional scopes: nationally 
in the U.S., in Washington state, and in the city Seattle, Washington. The data came from 
[Kaggle](https://www.kaggle.com/murderaccountability/homicide-reports?select=database.csv), which features 24 separate 
columns as an overview of details and 638454 cases.

* The Kaggle data was collected by the FBI’s Supplementary Homicide Report and Freedom of Information Act and compiled by 
the Murder Accountability Project. It is stated that the database contains homicides dating back to 1980 plus the 22,000 
not reported to the Justice Department, with classifications on race, sex, and age of both perpetrators and victims.

We wanted to explore the information through filtering data for details we would need. Things such as categorizations of 
the perpetrator and victim, weapons used, areas with higher homicide rates, span of years and number of murders happening, 
all of this information was used to answer our research queries.

### Summary Information
Here are some questions we thought of when planning our project:

* How do factors of the perpetrator and victim (ex. age, race, etc.) correlate?
* Are there trends between perpetrator age and homicide type?
* What are the comparisons of homicide rates between other states?
* Which cities of each state have the most homicide cases? Any concentrations?
* How do current homicide rates within this decade compare to rates from previous decades (ex. 2000-2010, 1990-2000)?

These questions branched into more and more questions as we sorted through our data. Some statistics we sought out for 
were:

* Average age of murder victims in Washington State
* Numbers of cities with lowest number of homicide cases
* Largest number of homicides in a single state
* Number of descriptions/categories (columns) in our data frame 
* Number of total cases (rows) altogether

#### Summary Table
The second set of questions were the questions we wanted to experiment with our data set. We decided to summarise the answers in our _Summary_Information.R_ file, which is displayed underneath, answering the questions in order (rendered in comments).

```{r echo = FALSE, message = FALSE}
source("Summary_Information.R", local = knitr::knit_global())
summary_list
```

#### Homicide Counts Over the Years
One point of interest we wanted to visualise was the amount of homicide cases in the United States over the years 1980 to 2018, by noticing any huge dips or spikes in the graph and general form. Starting from 1980, there is a gradual decline of homicide cases until 1985, gradually increasing until 1999 then returning to a longer slow decline. In 1999, we see the lowest point of the graph followed by the steepest rate of change leading to 2000, then homicide cases staying that high and even peaking in the years 2006 and 2010. The sheer number of homicide cases in the last two decades shadow over the amounts in the 1980s and 1990s, with the amount essentially doubling. From 2010, there is a noticeable decline of homicide cases before leading off with a small incline in recent years. 

```{r echo = FALSE}
source("Scatterplot.R", local = knitr::knit_global())
year_incident_plot
```

#### Distribution of Ages of Victims and Perpetrators

```{r, echo = FALSE, fig.width = 10, fig.height = 6, warning = FALSE}
source("histogram.R", local = knitr::knit_global())
grid.arrange(perp_hist, vict_hist, nrow = 1)
```

#### Popular Weapons Used in Homicide
We wanted to investigate if there was a more commonly used weapon within all homicide cases. What shows is the 
overwhelmingly distribution of weapon usage, especially the handgun as it is the top weapon used for homicide. The next 
common weapons are a firearm, knife, then blunt objects. The 

```{r, echo = FALSE, fig.width = 14, fig.height = 8}
source("barchart.r", local = knitr::knit_global())
weapon_barchart
```

#### Closing Thoughts
